---
title: "Day 3 Lecture"
author: "Cecilia Y. Sui"
date: "01/12/2022"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options)) # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines) == 1) { # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(more, x[lines], more)
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})
```


# Day 3 Outline:
1. Functions and Arguments 
2. Control Statements (if, if-else, for, while) and Random Sampling 


# 1. Functions
"Function is a set of instructions that performs a specific task, typically taking
input (known as arguments), processing these inputs, and returning an output."
Arguments is a value that you pass to a function when you call it. The function then uses this value as input to perform its operations. Think of it like giving ingredients to a chef; the ingredients (arguments) are what the chef (function) needs to make a meal (perform a task or calculation). 
R comes with many functions that you can use to do sophisticated tasks like
random sampling. For example, you can round a number with the **round()**
function, or calculate its factorial with the **factorial()** function. Using a
function is pretty simple: just write the name of the function and then the data
you want the function to operate on in parentheses. In fact, you have been using
functions in Day 2's problem set, including "sum(), mean(), max()" when analyzing
the dataframes.  

```{r}
# The following two functions come with base R.
# You do not need to install and load new packages to use them.
round(3.1415)
factorial(3)
```

The data that you pass into the function is called the function’s argument. The
argument can be raw data, an R object, or even the results of another R
function. In this last case, R will work **from the innermost function to the
outermost**.
```{r}
die <- 1:6
die
mean(1:6)
mean(die)
round(mean(die))
ceiling(mean(die)) # rounds up
floor(mean(die)) # rounds down
ceiling(4.000001)
floor(3.99999999)
```

```{r eval=FALSE}
round(3.1415926535, numbers = 1)
# Error in round(3.1415, numbers = 3) : unused argument (numbers = 3)
?round
round(x = 3.1415926535, digits = 3)
```



## 1.1 Sampling using the sample() function 

Base R comes with the sample() function, which takes two arguments: 

1. A vector named x 
2. A number named size

sample() will return the elements sampled from the vector. The return value is
**not** a list, which means if you need to do list operations, you need to first
convert it to a list.
```{r}
sample(x = 1:4, size = 2)
# sample(x = 1:4, size = 2, replace = TRUE)
die <- 1:6
sample(x = die, size = 3) # can input object

# Be careful with the scope of variables here.
# x is a local variable, NOT a global variable.
# This means that x can only be accessed within the function call sample()
```

Many R functions take multiple arguments that help them do their job. You can
give a function as many arguments as you like as long as you separate each
argument with a comma.

Every argument in every R function has an argument name. For example, x and size
are both argument names for the function sample().

You can specify which data should be assigned to which argument by setting a
name equal to data, as in the preceding code. This becomes important as you
begin to pass multiple arguments to the same function. Argument names help you
avoid passing the wrong data to the wrong argument.

However, using names is **optional**. You might notice that R users do not often
use the name of the first argument in a function. So you might see the previous
code written as:

```{r}
?sample
sample(x = die, size = 3) # best practice 
# or equivalently
sample(die, 3)
```

Often, the name of the first argument is not very descriptive, and it is usually
obvious what the first piece of data refers to anyways.

"But how do you know which argument names to use? If you try to use a name that a
function does not expect, you will likely get an error:"

If you are not sure which names to use with a function, you can look up the
function’s arguments with **args()**. To do this, simply plug in the name of the
function into the parentheses. For example, you can see that the round function
takes two arguments, one named x and one named digits:

```{r}
args(round)
round(3.14, digits = 1)
args(ceiling)
args(floor)
```

Did you notice that args shows that the digits argument of round is already set
to 0?

Frequently, an R function will take optional arguments like digits. These
arguments are considered optional because they come with a **default value**.
You can pass a new value to an optional argument if you want, and R will use the
default value if you do not.

For example, round() will round your number to 0 digits past the decimal point
by default. To override the default, you can always specify your own argument
value. For example,
```{r n}
round(3.1415)
round(3.1415, digits = 2)
```

When you first start to learn R, it is recommended that you write out the
**names of all your arguments** when you do a function call with multiple
arguments. This not only helps you to learn the functions and their arguments
better and faster, but also helps others to understand your code more easily.
Moreover, you might not always remember correctly how the arguments should be
ordered when first using these functions. Writing out the argument names helps
**prevent errors**.

After using R for a while, you might notice that it is often for R users to skip
the names for the first one or two arguments, and only write out the names for
the rest of arguments.

"If you do not write out the names of your arguments, R will match your values to
the arguments in your function by order." You can learn about the correct order
by **running the args() function, hovering over the function call, or using the
help pages **.

"For example, in the following code, the first value, die, will be matched to the
first argument of sample, which is named **x**. The next value, 1, will be
matched to the next argument, **size**":
```{r}
?sample
sample(die, 1)
```

As you provide more arguments, it becomes more likely that your order and R’s
order may **not align**. As a result, values may get passed to the wrong
argument. Argument names prevent this. R will always match a value to its
argument name, no matter where it appears in the order of arguments.

For example, if we reverse the order of arguments in the function call above, it
is still equivalent.
```{r}
sample(size = 3, x = die) # reversed order, specify argument name
sample(10)
```



### 1.1.1 Sampling without Replacement 

When we use the **sample()** function to do random sampling, the default is
sampling without replacement, where **replace = FALSE**.
```{r eval=FALSE}
?sample()
sample(x = die, size = 3)
# sample(x, size, replace = FALSE, prob = NULL)
```


### 1.1.2 Sampling with Replacement 

To do sampling with replacement, you can add the argument **replace = TRUE**:
```{r}
sample(x = die, size = 6, replace = TRUE)
sample(x = c(0, 1), size = 10, replace = TRUE) # 10 Bernoulli trials
```

The argument **replace = TRUE** causes sample() to sample **with replacement**.
As a result, sample() may select the same value on the second draw. Each value
has a chance of being selected each time. It is as if every draw were the first
draw.

"Sampling with replacement is an easy way to create independent random samples."
Each value in your sample will be a sample of size one that is independent of
the other values. For example, this would be how we simulate a pair of fair
dice:
```{r n}
die
sample(x = die, size = 2, replace = TRUE)
```


Congratulate yourself! You’ve just run your first simulation in R! 

You now have a method for simulating the result of rolling a pair of dice. If
you want to add up the dice, you can feed your result straight into the sum()
function:
```{r}
dice <- sample(die, size = 2, replace = T)
dice
# sum(dice)
```

What would happen if you call dice multiple times? Would R generate a new pair
of dice values each time? Let’s give it a try:
```{r}
dice
dice
dice
```

Nope. 

Each time you call dice, R will show you the result of that one time you called
sample and saved the output to dice. R won’t rerun **sample(die, 2, replace =
TRUE)** to create a new roll of the dice. This is a relief in a way. Once you
save a set of results to an R object, those results do not change. Programming
would be quite hard if the values of your objects changed each time you called
them.

However, it would be convenient to have an object that can **re-roll** the dice
whenever you call it. You can make such an object by writing your own **R
function** which we will cover in the following section.


## \textcolor{red}{In-class exercises 3.1:}

1. Sample with replacement from 1 to 1000 inclusive of size 20.
```{r}
set.seed(12345)
sample(1:100, 5)
```

2. Sample without replacement from 1 to 1000 inclusive of size 20.
3. What does the following code do?
```{r}
x <- 1:10
sample(x[x > 10])
```

4. We are going to use the world total fertility rate dataset again to do some
   sampling here. First, load your dataset as worldTFR.

5. Sample without replacement of size 100 from the TFR column in the dataset.
   What is the mean of your sample? Compare it to the mean of the entire column.

6. Repeat the sampling in Q5 five times, and report the means for each sampling.
   Compare the means to the mean of the entire column TFR.

7. Take the average of all the means obtained from Q6. What do you get? Is the
   new average closer to the true average for the column TFR?



## 1.2 Writing Your Own Functions 

To recap, you already have working R code that simulates rolling a pair of dice.

```{r}
die <- 1:6
dice <- sample(die, size = 2, replace = TRUE)
dice
sum(dice)
```

You can retype this code into the console anytime you want to re-roll your dice.
However, this is an awkward way to work with the code. It would be easier to use
your code if you wrapped it into its own **function**, which is exactly what we
will do now. You are going to write a function named roll that you can use to
roll two virtual dice. When you are finished, the function will work like this:
each time you call **roll()**, R will return the sum of rolling two dice:

```{r eval = FALSE}
# This code block is just for illustration.
# It is not executable yet, since we have not defined roll().
roll()
# 10
roll()
# 8
roll()
# 7
```

Functions might sound fancy when first hearing about it. They are just another
type of R object. Instead of containing data, like vectors, matrices, or
dataframes, functions contain code. This code is stored in a special format that
makes it easy to reuse the code in new situations throughout your program. You
can write your own functions by recreating the following format.


### 1.2.1 The Function Constructor

Every function in R has three basic parts: 

1. a function name
2. a body of code
3. a set of arguments

To make your own function, you need to replicate these parts and store them in
an R object, which you can do with the function **function()**. To do this, call
function() and follow it with a pair of braces, \{\}:

```{r}
my_function <- function( ) {
  ### the body of the function
  return( ) # the return value(s)
}
```


**function()** will build a function out of whatever R code you place between
the braces.

For example, you can turn your dice code into a function by running:
```{r}
roll <- function() {
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE)
  s <- sum(dice)
  return(s) # return value
}

roll()
```

Notice that I indented each line of code between the braces here. The
indentation is just for readability, which makes the code chunk easier for you
to read. It has **NO** impact on how the code executes. R ignores spaces and
line breaks inside these braces, and executes one complete expression at a time.

When writing the code between the braces, just hit the Enter key between each
line after the first brace, \{. R will wait for you to type the last brace, \},
before it responds. If you are using RStudio, when starting a new line between
braces, it automatically aligns with your code above.

Don’t forget to save the output of function to an R object. This object will
become your new function that you can call. To use it, write the object’s name
followed by an open and closed parenthesis. For example, if we call **roll()**
three times, it should output three different outputs.
```{r}
roll()
roll()
roll()
```

You can think of the parentheses as the “trigger” that causes R to run the
function. If you type in a function’s name without the parentheses, R will show
you the code that is stored inside the function. If you type in the name with
the parentheses, R will run that code:
```{r}
roll # view the code

roll() # run the function
```

The code that you place inside your function is known as the **body of the
function**. When you run a function in R, R will execute all of the code in the
body and then return the value you put in return().

**return()** is not required to write R functions. If you do not have a return
value specified, R will return the result of the last line of code inside the
braces.

If the last line of code does not return a value, neither will your function. So
if you want your function to have return values, you need to make sure that
either you final line of code returns a value or you specify the return value
inside your return().

One way to check this is to think about what would happen if you ran the body of
code line by line in the command line. Would R display a result after the last
line, or would it not?

Here’s some code that would display a result:

```{r eval = FALSE n}
dice
1 + 1
sqrt(2)
```

And here’s some code that would not:
```{r eval=FALSE n}
dice <- sample(die, size = 2, replace = TRUE)
two <- 1 + 1
a <- sqrt(2)
# df[1, ] <- 0
```

Do you notice the pattern? These lines of code do not return a value to the
command line; they save a value to an object. Generally, if the last line of
code in your function definition is an assignment statement, it will not return
anything.


### 1.2.2 Arguments

What if we removed one line of code from our function and changed the name die
to custom_die, like this?
```{r eval = FALSE n}
roll2 <- function() {
  dice <- sample(custom_die, size = 2, replace = TRUE)
  sum(dice)
}
```

Now we will get an error when we run the function. The function needs the object
custom_die to do its job, but there is no object named custom_die to be found in the
environment:
```{r eval = FALSE n}
roll2()
# Error in sample(custom_die, size = 2, replace = TRUE) :
# object 'custom_die' not found
```

You can provide custom_die when you call roll2 if you make custom_die an argument of the function. To do this, put the name custom_die in the parentheses that follow function when you define roll2:

"Say you want to make a different sided die"
```{r}
roll2 <- function(custom_die) {
  dice <- sample(custom_die, size = 2, replace = TRUE)
  return(sum(dice))
}
```

Now roll2 will work as long as you provide custom_die when you call the function
roll2(). You can take advantage of this to roll different types of dice each
time you call roll2.

Remember, we’re rolling pairs of dice that have different number of sides:
```{r}
roll2(custom_die = 1:4) # 4-sided
roll2(custom_die = 1:6) # 6-sided
roll2(custom_die = 1:20) # 20-sided
roll2(c(1, 2, 3, 4, 4, 6)) # without the argument name, it works the same
```

Notice that roll2 will still give an error if you do not specify a value for the
custom_die argument when you call roll2.
```{r eval = FALSE}
roll2()
# Error in sample(bones, size = 2, replace = TRUE) :
#  argument "custom_die" is missing, with no default
```

As mentioned in the Error Message, we can fix the error by giving the custom_die
argument a default value. To do this, set custom_die equal to a value when you define

```{r}
roll2 <- function(custom_die = 1:6) {
  dice <- sample(custom_die, size = 2, replace = TRUE)
  return(sum(dice))
}
```

Now you can provide a new value for custom_die if you like. If you do not provide any value for the custom_die argument, roll2() will use the default value that we defined above.

```{r}
roll2() # using custom_die = 1:6
roll2(c(1, 2, 2, 4, 4, 5)) # mario party custom_die
```

Again, **return()** is not strictly needed in R. R will return the last value
you calculate in the function by default, but **return()** makes it clearer.

```{r n}
roll3 <- function(custom_die = 1:6) {
  dice <- sample(custom_die, size = 2, replace = TRUE)
  sum(dice) # It is equivalent to just writing sum(dice)
}

roll3()
```

In R, the **return()** function can return only **one single object**. If we
want to return multiple values in R, we can use a list (or other objects) that
stores multiple values and return it as a single object.

```{r}
roll4 <- function(custom_die = 1:6) {
  dice <- sample(custom_die, size = 2, replace = TRUE)
  l <- list(sum(dice), dice)
  return(l)
}
roll4()
# The first item returned is sum(dice).
# The second item returned is the list dice that stores 2 values.
```



You can give your functions **as many arguments as you like**. Just list their
names, separated by commas, in the parentheses that follow function(). When the
function is run, R will replace each argument name in the function body with the
value that you provide for the argument.

* If you do not provide a value, R will replace the argument name with the
  argument’s default value if you defined one.
* If you forget to provide a value for an argument that does not have any
  default value defined, R will stop execution and print an Error Message asking
  for it!

To summarize, **function()** helps you construct your own R functions. 

1. Create a body of code for your function to run by writing code between the
   braces that follow **function()**.
2. Create arguments for your function to use by supplying their names in the
   parentheses that follow **function()**.
3. Give your function a name by saving its output to an R object.
4. Specify the return values if you want your function to return anything. 

Once you have created your function, R will treat it like every other function
in R, such as sum(), mean(), round(), etc. Think about how useful this is!

In the previous exericise, I asked you to sample the column TFR five times and
calculate their averages. What if I asked you to do it 1000 times? Without
functions, repeating the same chunk of code could be terrifying. As you learn to
program in R, you will be able to create new, customized, reproducible tools for
yourself whenever you like.


## \textcolor{red}{In-class exercises 3.2:}

1. Write a function called func_square() that takes one argument of type integer
   and returns the square of the number.

2. Write a function that prints out just your **first name**, but returns your
   **first and last name** together. Think about what should be your arguments
   here. Use **print()** function to do the printing.

3. Write a function called **my_sampling()** that samples without replacement
   from 1 to 100 inclusive of a given size. The function should take one
   argument that specifies the size of the returning vector.

4. Write a function that given a dataframe will return its column names.

5. Write a function that given a vector and an integer will return whether the
   integer is inside the vector.

6. Write a function that randomly samples from the TFR column in the worldTFR
   dataset. The users should be able to modify the size of the sample, and
   whether to sample with or without replacement.

7. Use the function you created in Q6 to do random sampling with replacement
   seven times. Calculate the mean for each sample. Then compute the average of
   the means. Compare it to the true mean.  





# 2. Control Statements
- imagine "traffic signals for a computer" 
In order to control the flow of execution of expressions in R, we make use of
the **control structures**. These control structures are sometimes called loops
in R. There are eight types of control structures in R (if/if-else, switch, for,
while, nested loops, repeat and break, next, return), but we will only focus on
**three here (if/if-else, for, while)**. Since all **while** loops can be
re-written as **for** loops, we will mainly work on **for** loops and briefly
introduce **while** loops.  

Whenever possible, it is more efficient to use **built-in functions** in R
rather than control structures. This facilitates the flow of execution to be
controlled inside a function. Control structures define the flow of the program.
The decision is then made after the variable is assessed.



## 2.1 $if$ Condition in R 

The **if** and **if-else** conditions in R enforce **conditional execution** of
the code. They are an important part of R's decision-making capability. It
allows us to make a decision based on the result of a condition. The **if**
statement contains a condition that evaluates to a logical output (TRUE = 1 /
FALSE = 0).

* It runs the enclosed code block if the condition evaluates to TRUE.
* It skips the code block if the condition evaluates to FALSE. 


```{r eval=FALSE}
# Syntax:
# This code block is not executable. Just for illustration.
if (test_expression) {
  statement
}
```

R is **NOT an indentation-based programming language**, so indentation or space
characters here would not make a difference as we illustrated above in the
functions section. You can put your braces at the end of the last line of your
code inside the loop, or start a new line just for the closing braces.

Let's do some examples! 

```{r}
my_cond1 <- TRUE
# my_cond <- T
if (my_cond1) {
  print("My condition is TRUE")
}

my_cond2 <- FALSE
if (my_cond2) {
  print("My condition is TRUE")
}


values <- 1:10
if (length(values) == 10) {
  print(values)
}

a <- 5
b <- 6
if (a < b) {
  print("a is smaller than b")
}
```


We will use a built-in dataset to illustrate how to recode variables using these
control statements.
```{r}
df <- longley
# You can use the help function to get more details on the dataset
# ?longley
# View(df)
```

For example, we can check to see if the year of the first row is 1947: 
```{r}
if (df$Year[1] == 1947) {
  print("Great, it's 1947!")
}
```

We can also check if there is any year where the number of unemployed exceeds
300.
```{r}
if (any(df$Unemployed > 300)) {
  print("Yes")
}
```


## 2.2 $if-else$ Condition in R 

We use the **else** statement with the **if** statement to enact a choice
between two alternatives. If the condition within the **if** statement evaluates
to FALSE, it runs the code within the **else** statement.

For example:
```{r}
a <- 88
b <- 100

if (a > b) {
  print("a is greater than b")
} else {
  print("b is greater than a")
}
```

We can use the **else if** statement to select between multiple options. For
example:
```{r}
a <- 100
b <- 88
if (a < b) {
  print("a is smaller than b")
} else if (a == b) {
  print("a is equal to b")
} else { # a > b
  print("a is greater than b")
} 
```

There is also an $ifelse()$ function in R. It acts like the **if-else**
structure. It is not very commonly used, but serves as a shortcut sometimes.

The following code is the syntax of the $ifelse()$ function in R:

```{r eval=F}
# This code block is not executable. Just for illustration.
ifelse(condition, exp_if_true, exp_if_false)
```


* condition is the condition that evaluates to either TRUE or FALSE
* exp_if_true is the expression that is returned if the condition results in
  TRUE
* exp_if_false is the expression that is returned if the condition results in
  FALSE

```{r}
a
ifelse(a < 7, "a is less than 7", "a is greater than 7")
```


We can use the **ifelse()** function to modify values in our dataframe. For
example, we can create a new column in the dataframe called PopOver100 where its
value is TRUE if the population is greater than 100 and FALSE otherwise.
```{r eval = FALSE}
df$PopMean <- ifelse(df$Population > mean(df$Population), TRUE, FALSE)
# df$PopMean2 <- df$Population > mean(df$Population)
View(df)
```

You might remember Q17 and Q18 from Day2's problem set. You can solve them using
this **ifelse()** function.

```{r eval = F n}
# Q17.
resources <- read.csv("resources.csv")

resources$oil[is.na(resources$oil)] = mean(resources$oil, na.rm = TRUE) 
resources$logGDPcp[is.na(resources$logGDPcp)] = mean(resources$logGDPcp, na.rm = TRUE) 
resources$regime[is.na(resources$regime)] = mean(resources$regime, na.rm = TRUE) 
resources$metal[is.na(resources$metal)] = mean(resources$metal, na.rm = TRUE) 
resources$illit[is.na(resources$illit)] = mean(resources$illit, na.rm = TRUE) 
resources$life[is.na(resources$life)] = mean(resources$life, na.rm = TRUE) 
summary(resources)

colnames(resources)[7] <- "life_expectancy"

# categorical variables: represents categories or group
# examples: Colors: Red, Blue, Green. Sizes: Small, Medium, Large. Types of Animals: Dogs, Cats, Birds. 

# as.factor create levels.
# without as.factor, "low" will be stored as character, with as factor "low" will be stored as vector of integer values with a corresponding set of character label, therefore treated as categorical variable. unordered factors by default.

resources$life <- as.factor(ifelse(resources$life_expectancy < 50, "low",
  ifelse(resources$life_expectancy < 70, "medium",
    ifelse(resources$life_expectancy < 100, "high")
  )
))
View(resources)

View(resources)
# Q18.
# resources$democracy <- ifelse(resources$regime >= 0, 1, 0)
```



## 2.3 $for$ Loop in R 

"The **for** loop in R, repeats through sequences to perform repeated tasks. They
work with an iterable variable to go through a sequence." The following code is
the syntax of **for** loops in R:

```{r eval = FALSE}
# This code is not executable. Just for illustration.
for (variable in sequence) {
  code_to_repeat
}
```

* variable is the iterative variable,
* sequence is the sequence which we need to loop through,
* code_to_repeat is the code that runs every iteration.


Here is an example: 
```{r}
vec <- c(2, 7)
vec
vec2 <- c(10:17)
vec2
for (i in vec) {
  print(vec2[i])
}
```


With the **for** loops, we can now iterate through the entire dataframe. 

For example, if you want to list all the years: 
```{r}
dim(df)
n1 <- dim(df)[1]
n1
# n2 <- dim(df)[2]
for (i in 1:16) {
  print(df$Year[i])
}
```

You can also recode the variables using **for** loops and the **if-else**
statement. We can start with listing out the years where the number of
unemployed exceeds 300.
```{r}
for (i in 1:dim(df)[1]) {
  if (df$Unemployed[i] > 300) {
    print(df$Year[i])
  }
}

# For good practice, you can assign dim(df)[1] to a new variable.
# For example:
L <- dim(df)[1]
```

Let's create a new column called ratio, and initialize it with NAs.  

```{r}
df$ratio <- NA
```

Now we want to recode the ratio variable, where we divide Employed over the sum
of Unemployed and Employed.
```{r}
# L <- dim(df)[1] # 16
for (i in 1:L) {
  df$ratio[i] <- df$Employed[i] / (df$Unemployed[i] + df$Employed[i])
}
View(df)
```

There is also a shortcut to get the ratio column.
```{r}
df$ratio2 <- df$Employed / (df$Unemployed + df$Employed)
```

## 2.4 $while$ Loop in R

**while** loops work similarly to **for** loops. The following code is the
syntax of a **while** loop.
```{r eval = FALSE}
while (test_expression) {
  code_to_repeat
}
```

* test_expression is evaluated to a logical output (TRUE / FALSE)
* If TRUE, code_to_repeat will be executed. If FALSE, R skips the code_to_repeat.

However, different from the test_expression in the **if-else** statement, in
**while** loops, the test_expression will be evaluated again when the execution
of code_to_repeat is completed. The process **repeats** each time **until
test_expression evaluates to FALSE**. When test_expression evaluates to FALSE,
the loop exits.

A trivial example: 
```{r}
while (FALSE) {
  print("F")
}
```
The code inside the **while** loop is never executed, because the
test_expression is always FALSE.

If we flip the FALSE to TRUE: 
```{r eval = FALSE}
# Don't run please
while (TRUE) {
  print("T")
}
```

This creates an endless **while** loop. It will keep running and printing "T"s
in your console.

You should **NEVER** set your test_expression to the logical TRUE directly when
writing programs. This could lead R to abort your session.

Here is another example: 
```{r}
i <- 1 # i is initialized to 1
while (i <= 6) {
  print(i)
  i <- i + 1 # increment i
}
```


Remember: **all for loops can be written as while loops, and vice-versa. **

Thus, when making a choice between the two, we should always consider which one
is more appropriate for the application. In general, you should use a **for**
loop whenever you already known how many times the loop should be executed (the
number of iterations). If the execution of the loop body is based on certain
condition, you should pick **while** loops instead. You will encounter **for**
loops a lot more than **while** loops in QPM I.

Just to illustrate the equivalency: 
```{r}
vec <- c(1:10)
for (i in vec) {
  print(vec[i])
}


vec <- c(1:10)
i <- 1
while (i <= length(vec)) {
  print(vec[i])
  i <- i + 1
}
```



## \textcolor{red}{In-class exercises 3.3:}

1. Write a function that samples with replacement from a given range with a
   given size. The user should also be able to choose whether to do sampling
   with or without replacement.

2. Write a **for** loop that does random sampling of size 1 without replacement,
   10 times from the range 1 to 100. Remember to print your randomly sampled
   number for each iteration. Use the function you created in Q1.

3. Convert your **for** loop into a **while** loop. 

4. Write a **for** loop that does random sampling of size 100 with replacement,
   10 times from 1 to 10000. Calculate the mean for each iteration and store the
   means in a new vector called "samples". This time you do not need to print
   the sampled numbers for each iteration. Simply print out the vector that
   contains all the means. Use the function you created in Q1.
```{r}
# To create a vector to store your values:
length <- 10
samples <- rep(NA, length)
```

5. Calculate the average of the means obtained in Q4. 
