---
title: "Day4 Problem Set"
author: "Cecilia Sui"
date: "1/13/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Day 4 Outline:
1. Types of Statistical Data (numerical, categorical, ordinal)
2. Probability distributions: rnorm / dnorm / pnorm / qnorm
3. Descriptive Statistics


# Probability Distributions

1. Suppose the height of males at a certain school is normally distributed with
   a mean of $\mu=168$ cm and a standard deviation of $\sigma = 9$ cm.
   Approximately what percentage of males at this school are **taller** than 175
   cm?
```{r}
pnorm(175, mean = 168, sd = 9, lower.tail = FALSE)
```

2. Suppose the weight of a certain species of dogs is normally distributed with
   a mean of $\mu=33$ lbs and a standard deviation of $\sigma = 5.5$ lbs.
   Approximately what percentage of this species of otters weight less than 28
   lbs?
```{r}
pnorm(28, mean = 33, sd = 5.5, lower.tail = TRUE)
```

3. Suppose the height of flowers in a certain region is normally distributed
   with a mean of $\mu=24$ cm and a standard deviation of $\sigma= 5$ cm.
   Approximately what percentage of flowers in this region are **between 18 and
   25 cm** tall?
```{r}
pnorm(25, mean = 24, sd = 5) - pnorm(18, mean = 24, sd = 5)
```

4. Find the Z-score of the 99th quantile of the standard normal distribution.
```{r}
qnorm(0.99)
```

5. Find the Z-score of the 95th quantile of the standard normal distribution.
```{r}
qnorm(0.95)
```

6. Generate a vector of 5 normally distributed random variables with mean = 10
   and sd = 2
```{r}
set.seed(12345)
rnorm(5, mean = 10, sd = 2)
```
7. Generate a vector of 100 normally distributed random variables with mean = 50
   and sd = 5
```{r}
rnorm(1000, mean = 50, sd = 5)
```

8. Generate a vector of 1000 normally distributed random variables with mean =
   50 and sd = 25
```{r}
rnorm(1000, mean = 50, sd = 25)
```
9. Generate two histograms using the **hist()** function to view these two
   distributions.
```{r}
hist(rnorm(100, mean = 50, sd = 25))
hist(rnorm(1000, mean = 50, sd = 25))
```


\newpage
# Descriptive Statistics

This section touches on how to run linear regression briefly. Don't worry if you
do not understand the meaning of every item in your summary table for your
model. We will learn more about how to interpret them in QPM I and II.

1. Install and load the **faraway** package. Load the **gavote** dataset. Study
   the dataset using the help() function. Create a new variable called
   undercount by calculating the percentage of ballots that were not counted
   into votes. What’s the range and quantiles of this new variable? Draw a
   histogram to illustrate the distribution of this variable.  
```{r}
# install.packages("faraway")
library("faraway")
df <- gavote
df$undercount <- (df$ballots - df$votes) / df$ballots
quantile(df$undercount)
hist(df$undercount)
```
2. Create a new variable **perGore** for the percentage of votes for Gore. Use
   **plot(col1, col2)** to create a scatter plot for the columns: **perGore**
   and **perAA**. What do you see? How can you interpret the plot?
```{r}
df$perGore <- (df$gore / df$votes)
plot(df$perAA, df$perGore)
# Positive correlation between perAA and perGore
```


3. Let’s run a linear regression with undercount as the response and perAA as
   the predictor. (Please refer back to the lecture notes for the example we
   did.) Summarize the regression results and describe what you see. Can you get
   the coefficients? Use **plot()** to visualize undercount and perAA.
```{r}
lm_3 <- lm(undercount ~ df$perAA, data = df)
summary(lm_3)
plot(df$undercount, df$perAA)
```
4. Run a linear regression with undercount as the response variable and both
   perGore and perAA as predictors. Summarize the results and describe what has
   changed. Can you guess why we observe such change?
```{r}
lm_4 <- lm(undercount ~ perAA + perGore, data = df)
summary(lm_4)
```