---
title: "Day 5 Pset"
author: "Cecilia Sui"
date: "1/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Day 5 Outline:
1. Packages and Help Pages 
2. Introduction to dplyr 
3. Introduction to ggplot2 
4. Tips on Using R Markdown


# dplyr Basics 
We will use the resources dataset to do some exercises with dplyr
```{r include=FALSE}
library(dplyr)
library(ggplot2)
df <- read.csv("/Users/river/Documents/GitHub/RCamp2023/Day5/Pset/resources.csv")
colnames(df)
```

1. Select the first 10 rows and last 3 rows. Output the rows together as the
   same dataframe.
```{r}
dim(df)
test <- df %>% slice(c(1:10, 4528:4530))
test
```


2. Select all rows where Year is 1951 and regime is non-negative.
```{r, eval=FALSE}
df %>% filter(year == 1971, regime >= 0)
```

3. Order the rows by Year in descending order. Break ties using the column oil in ascending order.
```{r, eval=FALSE}
df %>% arrange(desc(year), oil)
```

4. Select the 100th to 110th rows, and find the row with the highest value for
   metal.
```{r}
df %>% slice(100:110) %>%
    slice_max(metal)
```


5. Randomly select 10 rows with replacement, and standardize the columns oil and metal.
```{r}
test <- df %>%
  na.omit() %>% # This way we are only dealing with complete cases
  slice_sample(n = 10, replace = TRUE) %>%
  mutate(
    oil_sd = (oil - mean(oil)) / sd(oil),
    metal_sd = (metal - mean(metal)) / sd(metal)
  )
```


6. Select columns: cty_name, year, regime, oil, illit, and life and store them
   into a new dataframe.
```{r}
new_df <- df %>%
  select(cty_name, year, regime, oil, illit, life)
```

7. Add a new column named new_oil using where the values are the product of standardized oil and regime type.
```{r}
new_df <- new_df %>%
  mutate(new_oil = ((oil - mean(oil, na.rm = TRUE)) / sd(oil, na.rm = TRUE)) * ((regime - mean(regime, na.rm = TRUE)) / sd(regime, na.rm = TRUE)))
```

8. Calculate the **average** oil level for **each year** from 1966 to 1997. 
```{r}
vec <- c()
j <- 1
for (i in 1966:1997) {
  vec[j] <- mean(new_df$oil[new_df$year == i], na.rm = TRUE)
  j <- j + 1
}
vec

new_df %>%
  group_by(year) %>%
  summarise(oil = mean(oil, na.rm = TRUE))
```

9. Calculate the **median** metal value for **each** regime type from -10 to +10.
```{r}
df %>%
  group_by(regime) %>%
  summarise(metal = median(metal, na.rm = TRUE))
```

# ggplot2 Basics 

1. Find all the variables that negatively correlates with TFR from the
   **worldTFR** dataset. Can you offer a reasoning of the relationships?
```{r, eval=FALSE}
tfr <- read.csv("/Users/river/Documents/GitHub/RCamp2023/Day5/Lecture/worldTFR.csv") %>%
    na.omit()

# In the lm() function, if you put a . after the ~ the function
# assumes to use all other columns
lmTFR <- lm(TFR ~ ., data = tfr)

# Values with a negative coefficient here will have some sort of negative
# correlation with TFR
summary(lmTFR)
```

2. Pick two variables from Q1 and use the package ggplot2 to plot them and TFR
   separately, where the two variables you choose should be on your x-axis, and
   TFR should be on the y-axis. Add appropriate labels. Compare the two plots
   and explain the differences.
```{r}
tfr <- read.csv("/Users/river/Documents/GitHub/RCamp2023/Day5/Lecture/worldTFR.csv") %>%
  na.omit()

ggplot(tfr, aes(x = LifeExpB, y = TFR)) +
  geom_smooth(method = "lm") +
  geom_point() +
  labs(y = "Total Fertility Rate", x = "Life Expectancy")

ggplot(tfr, aes(x = Pop1564Female, y = TFR)) +
  # geom_smooth(method = "lm") +
  geom_point() +
  labs(y = "Total Fertility Rate", x = "Pop1564Female")
```


3. Create a new dataframe named df1, such that the dataframe has two columns:
   Year and avg_TFR. The column Year should list the years with no duplicates,
   and the column TFR should store the average TFR of all countries during that
   specific year. Use the package ggplot2 to visualize the relationship between
   Year and avg_TFR. Describe what you see.
```{r}
df1 <- tfr %>%
  group_by(Year) %>%
  summarise(avg_TFR = mean(TFR, na.rm = TRUE))

ggplot(df1, aes(x = Year, y = avg_TFR)) +
  # geom_smooth(method = "lm") +
  geom_point() +
  labs(y = "Total Fertility Rate", x = "Years")
```

